---
title: "Activity Report PA1"
author: "Michael Garcia"
date: "August 1, 2017"
output:
  html_notebook: default
  html_document: default
  pdf_document: default
  word_document: default
---
###### License and Copyright Notice
     Project: https://github.com/pimikeymike/reproducible-research
     Copyright 2017 Michael Garcia. All Rights Reserved.
     Inquiries mgar_datascience at protonmail.com
     
     Licensed under the Apache License, Version 2.0 (the "License");
     you may not use this file except in compliance with the License.
     You may obtain a copy of the License at
         http://www.apache.org/licenses/LICENSE-2.0
     
     Unless required by applicable law or agreed to in writing, software
     distributed under the License is distributed on an "AS IS" BASIS,
     WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
     See the License for the specific language governing permissions and
     limitations under the License.

#### 1) Read and Process Data

```{r include = TRUE, echo = TRUE}
library(data.table)

webURL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
destURL <- file.path(getwd(),paste("activity_monitoring_data",".zip",sep = ""))
localFileURL <- file.path(getwd(),paste("activity",".csv",sep = ""))
download.file(webURL,destfile = destURL)

unzip(destURL)
#activityData <- as.data.frame(read.csv(localFileURL))
activityData <- read.csv(localFileURL, quote = "\"",stringsAsFactors = FALSE)
activityData$steps <- as.numeric(activityData$steps)
activityData$interval <- as.numeric(activityData$interval)
activityData$date <- as.Date(activityData$date,format = "%Y-%m-%d")

```


#### 2) Histogram of the total number of steps taken each day




``` {r include = TRUE, echo = TRUE}
#   2) Histogram of the total number of steps taken each day
activityDataAg <- aggregate(activityData$steps,list(activityData$date),sum)
names(activityDataAg) <- c("date","totalSteps")
#activityDataAg <- activityDataAg[complete.cases(activityDataAg),]
hist(activityDataAg$totalSteps, xlab = "TotalSteps", breaks = 10)
```


#### 3) Mean and median number of steps taken each day

``` {r include = TRUE, echo = TRUE}
#3) Mean and median number of steps taken each day
stepSummary <- as.table(summary(activityDataAg$totalSteps))
stepsMedian <- stepSummary[3]
stepsMean <- stepSummary[4]
```
The total median steps are `r as.integer(stepsMedian)` and the mean for the total steps taken is `r as.integer(stepsMean)`


#### 4) Time series plot of the average number of steps taken

``` {r include = TRUE, echo = TRUE}
# 4) Time series plot of the average number of steps taken
# removed na's
actDataAg <- activityDataAg[complete.cases(activityDataAg),]
#activityTS= ts(rnorm(48), start=c(2293,6), frequency=1)
plot(actDataAg$date,actDataAg$totalSteps,type = "l", xlab = "date",ylab = "total steps")
```


#### 5) The 5-minute interval that, on average, contains the maximum number of steps

``` {r include = TRUE,echo = TRUE}
# 5   The 5-minute interval that, on average, contains the maximum number of steps
activity5min <- activityData[complete.cases(activityData),]
activity5min <- aggregate(activity5min$steps,list(activity5min$interval),mean)
names(activity5min) <- c("interval","avgsteps")
activity5min$avgsteps <- round(activity5min$avgsteps,1)
plot(activity5min$interval,activity5min$avgsteps,type = "l", xlab = "5min interval", ylab = "5min Avg Steps")
```
The 5 minutes interval with the most steps is `r activity5min[which(activity5min$avgsteps == max(activity5min$avgsteps)),"interval"]`. The average number of steps at this interval is `r activity5min[which(activity5min$avgsteps == max(activity5min$avgsteps)),"avgsteps"]`.


######  6 Code to describe and show a strategy for imputing missing datas

``` {r include = TRUE, echo = TRUE}
#stepSummary <- as.table(summary(activityDataAg$totalSteps))
#rowsNA <- nrow(activityData[!complete.cases(activityData),])
```
There are `r  nrow(activityData[!complete.cases(activityData),])` days that have NA's


``` {r include = TRUE, echo = TRUE}
#  6 Code to describe and show a strategy for imputing missing data

stepSummary <- as.table(summary(activityDataAg$totalSteps))

rowsNA <- nrow(activityData[!complete.cases(activityData),])



#library(zoo)
activityIntMerge <- merge(activityData,activity5min, by = "interval", all.x = TRUE)
activityIntMerge$steps <- ifelse(is.na(activityIntMerge$steps) == TRUE, activityIntMerge$avgsteps, activityIntMerge$steps)
activityIntMergeSum <- aggregate(activityIntMerge$steps,list(activityIntMerge$date),sum)
names(activityIntMergeSum) <- c("date","totalSteps")
activityIntMergeSum$date <-  as.Date(activityIntMergeSum$date,format = "%Y-%m-%d")
activityIntMergeSum$totalSteps <- as.numeric(activityIntMergeSum$totalSteps)
```

The strategy to fill na's included calculating the average steps per interval. Then NA's were replaced with the average value in respect to its interval. The mean after replacing the NA's is `r as.integer(round(NAFillStepsSummary[4],0))` and the median is `r as.integer(round(NAFillStepsSummary[3],0))`. The total daily steps and the average per inverval had no significant changes after filling the NA's, as can be seen on the charts.



###### 7) Histogram of the total number of steps taken each day after missing values are imputed

``` {r include = TRUE , echo = TRUE}
hist(activityIntMergeSum$totalSteps, main = "Total Steps NA filled with Interval Avg",xlab = "TotalSteps", breaks = 10 )

NAFillStepsSummary <- as.table(summary(activityIntMergeSum$totalSteps))
```


``` {r include = TRUE, echo = TRUE}
activityIntMergeAvg<- aggregate(activityIntMerge$steps,list(activityIntMerge$interval),mean)
names(activityIntMergeAvg) <- c("interval","avgSteps")
activityIntMergeAvg$avgSteps <- round(activityIntMergeAvg$avgSteps,1)
plot(activityIntMergeAvg$interval,activityIntMergeAvg$avgSteps,type = "l", main = "Interval after NA Filled",xlab = "5min interval", ylab = "5min Avg Steps")
```


###### 8) Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends


``` {r include = TRUE, echo = TRUE}
###### 8) Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

activityData$wkdy <- weekdays(activityData$date)
activityData$weektype <- ifelse((activityData$wkdy == "Saturday" | activityData$wkdy == "Sunday"), "Weekend", "Weekday" )
activityWk <- activityData[complete.cases(activityData),]
activityWkend <- aggregate(activityWk[which(activityWk$weektype =="Weekend"),"steps"]
                           , list(activityWk[which(activityWk$weektype =="Weekend"),"interval"]),mean)
names(activityWkend) <- c("interval","avgSteps")
activityWkend$interval <- as.integer(activityWkend$interval)
activityWkend$avgSteps <- round(as.numeric(activityWkend$avgSteps),1)

activityWkdy <- aggregate(activityWk[which(activityWk$weektype =="Weekday"),"steps"]
                           , list(activityWk[which(activityWk$weektype =="Weekday"),"interval"]),mean)
names(activityWkdy ) <- c("interval","avgSteps")
activityWkdy$interval <- as.integer(activityWkdy$interval)
activityWkdy$avgSteps <- round(as.numeric(activityWkdy$avgSteps),1)
par(mfrow=c(2,1))
#par(mfcol=c(2,1))
plot(activityWkdy$interval, activityWkdy$avgSteps, type = "l", xlab = "", ylab = "WkDay Avg Steps")
plot(activityWkend$interval, activityWkend$avgSteps, type = "l",xlab = "interval", ylab = "WkEnd Avg Steps")
mtext("Weekend vs. Weekday Avg Steps", side=3, outer=TRUE, line=-1) 
```

